Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-02-11 paper/letter.tex
affiliations
first plot, one paragraph, minor reference edit

\documentclass[a4paper, usenatbib, 12pt]{article}
\usepackage{subfig}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{cite}
\usepackage[top=3.5cm, bottom=2.5cm, left=3.5cm, right=3.5cm]{geometry}

@@

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Lyman-alpha emission reveals an unusual fastly rotating
  compact dwarf galaxy}  
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$,
\\
\\
\scriptsize
{$^1$ Universidad de los Andes
\\
$^2$ Arizona
\normalsize
\\
\\
\textbf{
Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected pristine
conditions of the first galaxies.  
Until these early galaxy generations are observationally
detected, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.  
Here we report on one of such CDGs, \tol, which presents
features in its Lyman-$\alpha$ emission thad had evaded theoretical
interpretation so far. 
We show that these special features, a symmetric triple peaked
emission line, are naturally explained by gas rotation.  
We find that the Lyman-$\alpha$ emission region in \tol\ should
have a rotational velocity of $V_{r}=300$ km s$^{-1}$ and a neutral
Hydrogen column density of $\log N_{HI} / \mathrm{atoms\ cm}^{-2} =
  20.5$.   
Using archival observational information we find that
the diameter for that region should be in the range of
$110$ pc $<D<$ $340$ pc and its total dynamical mass 
should be between $2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times
10^{9}$M$_{\odot}$. 
This dynamical mass is at least $16\pm 9$ times
larger than the neutral mass hydrogen.
We argue that a possibility to explain the excess in
dynamical mass is the presence of a super-massive black hole. }

@@

3. [Observations require interpretation, common features and
  models]. The Lyman-$\alpha$ line is resonant. A yman-$\alpha$ photon
will follow a difussion-like process before escaping the galaxy or
being absorbed by dust. This makes it hard to interpret Lyman-alpha
line observations as the reulting profile is sensitive to the
dynamical, chemical and thermal conditions in the interstellar
medium. State-of-the-art models use mon

@@

Figure 1. shows the observational data for \tol with the overplot from
our best fit model from the full radiative transfer simulation. 
The parameters for the best fit are
$v_{max}=300$\kms, $\tau=1\times10^7$, $T=1.5\times 10^{4}$K and a
viewing angle $\theta<30$ degrees.

Figura Observed line + fit.

@@

8. [Implications for outflow+rotation in existing samples.]

@@

%=====

@@

The receeding velocity is $7785\pm 50$km s$^{-1}$, which translates
into a distance of $106.6$ Mpc (Hubble constant 73 Mpc km$^{-1}$
s$^{1}$)
The metallicity is $\sim Z_{\odot}/24$ \cite{Izotov04} as derived from optical
spectroscopy.

>>>>

\documentclass[a4paper, usenatbib, 12pt]{article}
\usepackage{subfig}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{cite}
\usepackage[icelandic,spanish,english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[top=3.5cm, bottom=2.5cm, left=3.5cm,right=3.5cm]{geometry}

@@

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Þú ert jörðin}
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$, MD$^3$
\\
\\
\scriptsize
{$^1$ Bogota
\\
$^2$ Tucson
\\
$^3$ Oslo
\normalsize
\\
\\
\textbf{
Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected pristine conditions of the first galaxies.  
Until these early galaxy generations are observationally
detected, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.  
Here we report on one of such CDGs, \tol, which presents
features in its Lyman-$\alpha$ emission thad had evaded theoretical
interpretation so far. 
We show that these special features, a symmetric triple peaked
emission line, are naturally explained by gas rotation.  
We find that the Lyman-$\alpha$ emission region in \tol\ should
have a rotational velocity of $V_{r}=300$ km s$^{-1}$ and a neutral
Hydrogen column density of $\log N_{HI} / \mathrm{atoms\ cm}^{-2} =
  20.5$.   
Using archival observational information we find that
the diameter for that region should be in the range of
$110$ pc $<D<$ $340$ pc and its total dynamical mass 
should be between $2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times
10^{9}$M$_{\odot}$. 
This dynamical mass is at least $16\pm 9$ times
larger than the neutral mass hydrogen.
We argue that a possibility to explain the excess in
dynamical mass is the presence of a super-massive black hole. }

@@

3. [Observations require interpretation, common features and
  models]. The Lyman-$\alpha$ line is resonant. A yman-$\alpha$ photon
will follow a difussion-like process before escaping the galaxy or
being absorbed by dust. This makes it hard to interpret Lyman-alpha
line observations as the reulting profile is sensitive to the
dynamical, chemical and thermal conditions in the interstellar
medium. State-of-the-art models use mon.

@@

Figure 1. shows the observational data for \tol

with the overplot from
our best fit model from the full radiative transfer simulation. 
The parameters for the best fit are
$v_{max}=300$\kms, $\tau=1\times10^7$, $T=1.5\times 10^{4}$K and a
viewing angle $\theta<30$ degrees.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{tolfit.png}
\caption{Nice fit}
\end{center}
\end{figure}

@@

8. [Implications for outflow+rotation in existing samples.] Why are
not more examples of triple peaked emission? 
The ubiuquity of galaxy outflows overimposed to the rotation
transforms the line into the common double or single peaked
observations.

@@

@@

\tol receeding velocity is $7785\pm 50$km s$^{-1}$, which translates
into a distance of $106.6$ Mpc (Hubble constant 73 Mpc km$^{-1}$
s$^{1}$)
Its metallicity is $\sim Z_{\odot}/24$ \cite{Izotov04} as derived from optical
spectroscopy.
